---
title: 'FINAL PROJECT'
output: html_notebook
---

This study allows us to revisit/renew

1. Regression modeling
2. Properties of Least Squares/Fitting "a line"
3. Multiple observation

Datasets for this study are

1. The main file: gauge.txt
2. Supplementary large-scale files: download the following folder Full Resolution Data.zip More information about the supplementary file can be found at http://iabp.apl.washington.edu/data.html as well as http://nsidc.org/data/G00791


## Question
The aim of this lab is to provide a simple procedure for converting gain into density when the gauge is in operation. Keep in mind that the experiment was conducted by varying density and measuring the response in gain, but when the gauge is ultimately in use, the snow-pack density is to be estimated from the measured gain.


## Scenario 1
```{r}
#df[sample(nrow(df), 10000), ]
#install.packages('L1pack')
#install.packages('quantreg')
#install.packages('ggplot2')
#library(L1pack)  # Used for least absolute deviations regression line
#library(quantreg)  # Used for quantile regression line
#library(ggplot2)
```
## Scenario 2
```{r}
library("dplyr")
Donations <- read.csv("io/Donations.csv")
Donors <- read.csv("io/Donors.csv")
Donations <- Donations[,c('Donor.ID', 'Donation.Amount')]
Donors <- Donors[,c('Donor.ID', 'Donor.Is.Teacher')]
df_merged <- merge(Donations,Donors,by="Donor.ID")
teacher <- df_merged[which(df_merged$Donor.Is.Teacher== 'Yes'),]
nonteacher <- df_merged[which(df_merged$Donor.Is.Teacher=='No'),]
```

```{r}
#true.fraction <- as.numeric(table(Donors$Donor.Is.Teacher)[2])
#print(paste0("The true fraction of donors that are teachers is ", true.fraction))
#sample <- df_merged[sample(nrow(Donations),10000),]
teacher_money_amount <- teacher$Donation.Amount
nonteacher_money_amount <- nonteacher$Donation.Amount
```

```{r}
# two-sided test
# null: true shift is zero, teacher donates the same as non-teachers
wilcox.test(teacher_money_amount, nonteacher_money_amount, paired = FALSE, correct = FALSE)
```

```{r}
# one-sided test- test for difference less than 0
# null: true shift is less than 0, teacher donates less than non-teachers
wilcox.test(teacher_money_amount, nonteacher_money_amount, paired = FALSE, correct = FALSE, alternative = "greater")
```

```{r}
# one-sided test- test for difference greater than 0
# null: true shift is greater than 0 teacher donates more than non-teachers
wilcox.test(teacher_money_amount, nonteacher_money_amount, paired = FALSE, correct = FALSE, alternative = "less")
```

## Scenario 3
```{r}

```

## Scenario 4
```{r}

```

## Scenario 4
```{r}

```