---
title: 'FINAL PROJECT'
output: html_notebook
---

## Question
The goal of this project is to build a recommendation system helping donors to connect to the projects they are interested to donate. We want to pair up donors to the classroom requests that will most motivate them to make an additional gift. If DonorsChoose.org can motivate even a fraction of those donors to make another donation, that could have a huge impact on the number of classroom requests fulfilled.


## Install packages
```{r}
#install.packages("mapproj")
#install.packages("ggmap")
#install.packages("DeducerSpatial")
#install.packages('L1pack')
#install.packages('quantreg')
#install.packages('ggplot2')
#library(L1pack)  # Used for least absolute deviations regression line
#library(quantreg)  # Used for quantile regression line
#library(ggplot2)
library("dplyr")
```

## Import Dataset
```{r}
#Donations <- read.csv("io/Donations.csv")
#Donors <- read.csv("io/Donors.csv")
#projects <- read.csv("io/projects_mini.csv")
#schools <-  read.csv("io/Schools.csv")[,c(1,3,4)]
#teachers <- read.csv("io/Teachers.csv")

# Merge Data Frames
# projects <- read.csv("io/Projects.csv")
# projects <- projects[-c(6,7,8,9)]
# write.csv(projects, "io/projects_mini.csv")
# df_merged <- merge(Donations,Donors,by="Donor.ID")
# df_merged <- df_merged[-3]
# df_merged <- merge(df_merged,projects,by="Project.ID")
# df_merged <- merge(df_merged,schools,by="School.ID")
# write.csv(df_merged, "io/df_merged.csv")

#merged <-  aggregate(list(donations=Donations$Donation.Amount), by=list(Project.ID=Donations$Project.ID), FUN= sum)
#merged <- merge(merged,projects, by = "Project.ID")
#merged <- merge(merged,schools, by = "School.ID")
#write.csv(merged, "io/merged_projects.csv")

# Merged dataframe based on Donation.ID
df_merged <- read.csv("io/df_merged.csv")

# Merged dataframe based on each Project.ID
merged <- read.csv("io/merged_projects.csv")
```


## Scenario 1
Question:Is there a seasonal pattern of donations? Is the total amount of received donation related to donation date? Is the total amount of received donation related to day of the week?
```{r}

```


## Scenario 2
Question: Test whether the teachers’ donation pattern is different from non-teachers’ donation pattern.(Test whether the fraction of teacher in the whole population is larger than the proportion of the donors that are non teacher)

```{r}
teacher <- df_merged[which(df_merged$Donor.Is.Teacher== 'Yes'),]
nonteacher <- df_merged[which(df_merged$Donor.Is.Teacher=='No'),]
teacher_money_amount <- teacher$Donation.Amount
nonteacher_money_amount <- nonteacher$Donation.Amount
```

2.1 Wilcoxon test- test for the difference in mean
```{r}
# one-sided test- test for difference greater than or equal to 0
# null: true shift is greater than 0 teacher donates more than non-teachers
wilcox.test(teacher_money_amount, nonteacher_money_amount, paired = FALSE, correct = FALSE, alternative = "less")
```

2.2 True fraction of teacher donors in the donor population
```{r}
true.fraction <- round(as.numeric(table(Donors$Donor.Is.Teacher)[2])/nrow(Donors)*100, 3)
print(paste0("The true fraction of teacher donors is ", true.fraction, "%"))
```


## Scenario 3
Question: Which state has the highest donor to population ratio? Is the amount of donation correlated with the region donors live in?
```{r}

```


## Scenario 4
Predict donors' donating behaviors to figure out how much a donor would donate for a project. 
4.1 In total, find out the states that have the most donations.
```{r}
# Figure out the rank of total donation amount for each state
state.total <-  aggregate(list(donations=df_merged$Donation.Amount), by=list(states=df_merged$Donor.State), FUN= sum)
state.total <-  state.total[order(state.total$donations, decreasing = TRUE),]
summary(state.total$donations)
```

4.2 On average, find out the states that have the most donation per transaction.
```{r}
# Figure out the rank of average donation amount per transaction for each state
state.avg <-  aggregate(list(donations=df_merged$Donation.Amount), by=list(states=df_merged$Donor.State), FUN= mean)
state.avg <- state.avg[order(state.avg$donations, decreasing = TRUE),]
summary(state.avg$donations)
```

4.3 Least Square Regression on predicting how much a donor will donate 
```{r}
sample <- merged[sample(nrow(merged), 100000), ]
sample=na.omit(sample)
ols = lm(formula = donations ~ Project.Type + Project.Grade.Level.Category+ Project.Cost+ School.Metro.Type + School.Percentage.Free.Lunch  , data = sample)
summary(ols)
```

4.4 Ridge Regression on predicting how much a donor will donate 
```{r}
# Ridge
#install.packages("MASS")
library(MASS)
ridge <- lm.ridge (donations ~ Project.Type + Project.Cost + School.Metro.Type + School.Percentage.Free.Lunch, data = sample, lambda = seq(0, 50, .001))
plot(ridge)
select(ridge)
```
```{r}
# install.packages("ISLR")
# install.packages("glmnet")
library("glmnet")
library(ISLR)
set.seed(2018)

sample = sample[c(4,7,10,11,17,18)]
x=model.matrix(donations~.-1,data=sample) 
y=sample$donations
fit.ridge=glmnet(x,y,alpha=0)
plot(fit.ridge,xvar="lambda",label=TRUE)
cv.ridge=cv.glmnet(x,y,alpha=0)
plot(cv.ridge)
```

4.5 Lasso Regression on predicting how much a donor will donate 
```{r}
fit.lasso=glmnet(x,y,alpha=1)
plot(fit.lasso,xvar="lambda",label=TRUE)
cv.lasso=cv.glmnet(x,y)
plot(cv.lasso)
coef(cv.lasso)
n <- nrow(x)
train <- sample(1:n, size = floor(0.6*n), replace = FALSE) # use 60% data for training
lasso.tr=glmnet(x[train,],y[train], alpha = 1)
lasso.tr
pred=predict(lasso.tr,x[-train,])
# calculate the mse
rmse= sqrt(apply((y[-train]-pred)^2,2,mean))
plot(log(lasso.tr$lambda),rmse,type="b",xlab="Log(lambda)")
lam.best=lasso.tr$lambda[order(rmse)[1]]
lam.best
coef(lasso.tr,s=lam.best)
```


## Scenario 5
Forests/ Bagging(Include Cross Validation)/ clustering - Predict how likely the donor is going to donate. 

5.1 Random Forests
```{r}
#install.packages('randomForest')
library(randomForest)

n <- nrow(sample)
train <- sample(1:n, size = floor(0.75*n), replace = FALSE) # use 75% data for training
rf.fit <- randomForest(donations ~ ., data = sample[train, ], importance = TRUE, ntree = 1000, nodesize = 10)
varImpPlot(rf.fit)
rf.pred <- predict(rf.fit, sample[-train, ], type = 'response')
# MSE
print(mean((rf.pred - sample[-train, ]$donations)^2))
print(mean(abs(rf.pred - sample[-train, ]$donations)))
print(mean(abs(rf.pred - sample[-train, ]$donations)/sample[-train, ]$donations))
```

5.2 K-Mean Clustering
```{r}
library(ggplot2)
ggplot(sample, aes(School.Percentage.Free.Lunch,donations, color = School.Metro.Type)) + geom_point()
set.seed(314)
irisCluster <- kmeans(iris[, c(1,3)], 3, nstart = 20)
irisCluster
```