---
title: "Math 189/289 Lab 10"
author: "Alex Hanbo Li and Jiaqi Guo"
output:
  html_notebook: default
  pdf_document: default
---

We are going to talk about ridge regression and lasso regression today.

# Ridge regression
## Theory
One way to control the complexity of a model is to penalize its magnitude. For example, in a linear regression problem
$$
\min_{\beta \in \mathbb{R}^p} \sum_{i=1}^n (y_i - x_i^\top \beta)^2,
$$
we can control the magnitude of coefficients $\beta$. Of course the magnitude of $\beta$ could be defined in different ways, e.g. two-norm $\|\beta\|_2$, one-norm $\|\beta\|_1$ and infinity-norm $\|\beta\|_{\infty}$. The ridge regression involves the two-norm penalty:
$$
\min_{\beta \in \mathbb{R}^p} \sum_{i=1}^n (y_i - x_i^\top \beta)^2 + \lambda \|\beta\|_2^2
$$
where $\lambda$ is a tuning parameter controling the level of regularization. Denote $X$ to be the $n$ by $p$ design matrix with rows to be $x_i^\top$, and $Y$ be the $n$ by 1 vector of $y_i$'s. Assume $X^\top X + \lambda I$ is invertible, we have an explicit solution to the ridge regression problem
$$
\hat \beta_{ridge} = (X^\top X + \lambda I)^{-1}X^\top Y.
$$
Recall that the solution to the ordinary least square regression is (assuming invertibility of $X^\top X$)
$$
\hat \beta_{ols} = (X^\top X)^{-1}X^\top Y.
$$
Two facts: When $\lambda \to 0$, $\hat \beta_{ridge} \to \hat \beta_{ols}$; when $\lambda \to \infty$, $\hat \beta_{ridge} \to 0$.

In the special case when $X$ is orthonormal (i.e. the columns of $X$ are orthonormal), we have
$$
\hat \beta_{ridge} = \frac{\hat \beta_{ols}}{1 + \lambda}.
$$
So we can see ridge estimator has an extra $1/(1 + \lambda)$ shrinkage factor. Therefore, there is bias on the ridge estimator.

## Fit the model
First, let's look at the package `MASS` for ridge regression.
```{r}
library(MASS)

# Generating the data
set.seed(20170315)
N <- 100      # Sample size

x1 <- runif(n=N)
x2 <- runif(n=N)
x3 <- runif(n=N)
x3c <- 10*x1 + x3 # New variable
ep <- rnorm(n=N)
y <- x1 + x2 + ep 
data <- cbind.data.frame(x1,x2,x3,x3c,y)

# OLS fit of 3-variable model using independent x3
ols <- lm(y~ x1 + x2 + x3, data = data)
summary(ols)


# OLS fit of 3-variable model using correlated x3.
olsc <- lm(y~ x1 + x2 + x3c, data = data)
summary(olsc)

# Ridge regression using independent variables
ridge <- lm.ridge (y ~ x1+x2+x3, data = data, lambda = seq(0, 50, .001))
plot(ridge)
select(ridge)

# Ridge regression using correlated variables
ridgec <- lm.ridge (y ~ x1+x2+x3c, data = data, lambda = seq(0, 50, .001))
plot(ridgec)
select(ridgec)
```

A better package is `glmnet`.
```{r}
# install.packages("ISLR")
# install.packages("glmnet")
library(ISLR)
library(glmnet)

Hitters=na.omit(Hitters)
x=model.matrix(Salary~.-1,data=Hitters) 
y=Hitters$Salary
```

The ridge-regression model is fitted by calling the glmnet function with $\alpha = 0$. It makes a plot as a function of log of lambda, and is plotting the coefficients.
```{r}
fit.ridge=glmnet(x,y,alpha=0)
plot(fit.ridge,xvar="lambda",label=TRUE)
```

## Model selection
We will do cross-validation to select the best $\lambda$.
```{r}
cv.ridge=cv.glmnet(x,y,alpha=0)
plot(cv.ridge)
```
The left vertical line is the one with the minimum MSE. The right vertical line is within one standard error of the minimum, which is a slightly more restricted model that does almost as well as the minimum. We will go for the second one to prevent overfitting.

At the top of the plot, you actually see how many non-zero variables coefficients are in the model. In this case, all 20 variables are included in the model (19 variables plus the intercept) and no coefficient is zero.

# Lasso
Instead of $L_2$ regularization, LASSO uses $L_1$ penalization, that is,
$$
\min_{\beta \in \mathbb{R}^p} \sum_{i=1}^n (y_i - x_i^\top \beta)^2 + \lambda \|\beta\|_1. 
$$
Because of the nature of $L_1$ norm, LASSO tends to give more sparse solution than ridge regression. This is typically useful in high-dimensional setting when the true model is actually a low-dimensional embedding.

## Fit the model
To fit the lasso model, you can specify $\alpha = 1$ to the fitting function.
```{r}
fit.lasso=glmnet(x,y,alpha=1)
plot(fit.lasso,xvar="lambda",label=TRUE)
```

## Cross validation to choose $\lambda$
```{r}
cv.lasso=cv.glmnet(x,y)
plot(cv.lasso)
cv.lasso$lambda.min
cv.lasso$lambda.1se
```
We can extract the coefficients.
```{r}
coef(cv.lasso)
```

```{r}
n <- nrow(x)
train <- sample(1:n, size = floor(0.6*n), replace = FALSE) # use 60% data for training
lasso.tr=glmnet(x[train,],y[train], alpha = 1)
lasso.tr
pred=predict(lasso.tr,x[-train,])

# calculate the mse
rmse= sqrt(apply((y[-train]-pred)^2,2,mean))
plot(log(lasso.tr$lambda),rmse,type="b",xlab="Log(lambda)")
lam.best=lasso.tr$lambda[order(rmse)[1]]
lam.best
coef(lasso.tr,s=lam.best)
```

# Random forest
Random forest is an ensemble of decision trees. It's kind of majority vote. Since the formulas for building a single decision tree are the same every time, some source of randomness is required to make these trees different from one another. Random Forests do this in two ways:

## Bagging: bootstrap aggregating
Bagging takes a randomized sample of the rows in your training set, with replacement.

## Sampling the variables
Instead of looking at the entire pool of available variables, Random Forests take only a subset of them.

Through these two sources of randomness, the ensemble contains a collection of totally unique trees which all make their classifications differently.

```{r}
# install.packages('randomForest')
library(randomForest)

n <- nrow(Hitters)
train <- sample(1:n, size = floor(0.75*n), replace = FALSE) # use 75% data for training
rf.fit <- randomForest(Salary ~ ., data = Hitters[train, ], importance = TRUE, ntree = 1000, nodesize = 10)
```
The `importance=TRUE` argument allows us to inspect variable importance as we’ll see, and the `ntree` argument specifies how many trees we want to grow.
```{r}
varImpPlot(rf.fit)
```
From the variable importance plot, *CHits*, *CAtBat*, *CRuns*, and *CRBI* are the most important factors for a player's salary.

Now we can use the random forest model to make predictions.
```{r}
rf.pred <- predict(rf.fit, Hitters[-train, ], type = 'response')
# MSE
mean((rf.pred - Hitters[-train, ]$Salary)^2)
# MAD
mean(abs(rf.pred - Hitters[-train, ]$Salary))
# relative
mean(abs(rf.pred - Hitters[-train, ]$Salary)/Hitters[-train, ]$Salary)
```

# K-Means clustering
As a clustering method, K-Means tries to cluster data points based on their similarity. The algorithm requires only one parameter, the number of clusters $k$. Formally, given observations $x_1, x_2, \cdots, x_n$, k-means clusters the $n$ observations into $k$ sets $S_1, S_2, \cdots, S_k$. The minimization problem is
$$ \min_{S_1, S_2, \cdots, S_k} \sum_{j=1}^k \sum_{\{i|x_i \in S_j\}} \|x_i-\mu_j\|^2, $$
where $\mu_j$ is the centroid of set $S_j$. The algorithm initializes with $k$ centroids, and labels each observation by the closest centroid. Then the algorithm iterates between two steps: update centroids within clusters and reassign each data point to the cluster whose centroid is the closest.

The following is an example with iris dataset.
```{r}
library(datasets)
head(iris)
```
Here is a visualization of the iris dataset.
```{r}
library(ggplot2)
ggplot(iris, aes(Sepal.Length, Petal.Length, color = Species)) + geom_point()
```
Now we apply k-means clustering.
```{r}
set.seed(314)
irisCluster <- kmeans(iris[, c(1,3)], 3, nstart = 20)
irisCluster
```
Let's take a look at how the clustering algorithm performs.
```{r}
table(irisCluster$cluster, iris$Species)
```
```{r}
irisCluster$cluster <- as.factor(irisCluster$cluster)
ggplot(iris, aes(Sepal.Length, Petal.Length, color = irisCluster$cluster)) + geom_point()
```
