---
title: "RentalListingsFinal"
author: "Justin Glommen"
date: "3/12/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
install.packages("jsonlite", repos="http://cran.r-project.org")
```

# Data

The dataset is provided by Two Sigma's Kaggle competition page. The dataset is in the JSON format, and so we must read it in for observation and manipulation.
```{r}
library('jsonlite')
trainData <- fromJSON("train.json", simplifyDataFrame=TRUE)
```

## Categorization

As the data is inputted via a JSON object, we must dissect it in order to utilize the numerical data.
```{r}
x <- 1
trainBathrooms <- matrix(data = NaN, nrow = length(trainData$bathrooms), ncol = 1)
trainPrices <- matrix(data = NaN, nrow = length(trainData$price), ncol = 1)
trainInterests <- matrix(data = NaN, nrow = length(trainData$interest_level), ncol = 1)
for (i in names(trainData$bathrooms)) {
  trainBathrooms[x,1] <- trainData$bathrooms[[i]]
  x <- x + 1
}

x <- 1
for (i in names(trainData$price)) {
  trainPrices[x,1] <- trainData$price[[i]]
  x <- x + 1
}

x <- 1
for (i in names(trainData$interest_level)) {
  trainInterests[x,1] <- trainData$interest_level[[i]]
  x <- x + 1
}
#hist(trainData$bathrooms[,1])
#matrix <- data.frame(x1= trainData$bathrooms, x2= trainData$price)
#correlation <- cor(matrix)
#library("corrplot")
#corrplot(correlation, method = "circle") #plot matrix
```



# Visualization
After initializing and sorting the individual categories properly into useable formats, we want to begin the analysis and visualize the data.
```{r}
# Graph the train prices
hist(trainPrices, breaks = 10000, xlim = c(0,11000))

# Plot number of bathrooms and how it correlates to prices
plot(trainBathrooms, trainPrices, ylim = c(0,11000))

# Plot the categorized interest_levels and their frequencies
plot(table(trainInterests))

# Gather the high interest houses, plot the prices
highInterest <- array(which(trainData$interest_level == "high"))
mediumInterest <- array(which(trainData$interest_level == "medium"))
lowInterest <- array(which(trainData$interest_level == "low"))

hist(trainPrices[highInterest], breaks = 250, xlim = c(0,11000))
hist(trainPrices[mediumInterest], breaks = 50, xlim = c(0,11000))
hist(trainPrices[lowInterest], breaks = 10000, xlim = c(0,11000))

# Plot the high interest houses versus the general interest houses prices
generalPrices <- trainPrices
highPrices <- trainPrices[highInterest]


mean(trainPrices)
mean(trainPrices[highInterest])
mean(trainPrices[mediumInterest])
mean(trainPrices[lowInterest])
```

