---
title: "Math189Pr_3"
output: html_document
---

Overall Question: How do we find clusters of palindromes? How do we determine whether a cluster is just a chance occurrence or a potential replication site?

# Import Dataset
```{r}
location <- read.table("hcmv-25kgjn1-1rfrtkc.txt", header = TRUE)
raw <- read.csv("RAW_DATA-2iwcznn-2kr2xw0.csv", header = TRUE)
```

# Random Scatter
Generate structure that examine the locations of the palindromes. Simulate 296 palindrome sites chosen at random along a DNA sequence of 229,354 bases using a pseudo random number generator. 
```{r}
set.seed(2017)
# Generate simple random sample
N <- 229354
n <- 296
theo <- sample.int( N, size = n)
theo
```
```{r}
# Visualization
hist(theo, breaks = 100, probability = TRUE, col = 3, main = "Simple Ramdom Sample Distribution")
lines(density(theo, adjust = 2), col = 2)
hist(location$location, breaks = 100, probability = TRUE, col = 3, main = 'Actual Data Distribution')
lines(density(location$location, adjust = 2), col = 2)
```

```{r}
# Group the data (Still Figuring out)
k <- 
tab <- table(cut(theo, breaks = seq(0, 23, length.out = k+1), include.lowest = TRUE))
head(tab, 10)
counts <- as.vector(tab)
head(counts, 10)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
