---
title: "Math189Pr_3"
output: html_document
---

Overall Question: How do we find clusters of palindromes? How do we determine whether a cluster is just a chance occurrence or a potential replication site?

# Scenario 1: Random Scatter
# Import Dataset
```{r}
location <- read.table("hcmv-25kgjn1-1rfrtkc.txt", header = TRUE)
data <- read.csv("RAW_DATA-2iwcznn-2kr2xw0.csv", header = TRUE)
```
Generate structure that examine the locations of the palindromes. Simulate 296 palindrome sites chosen at random along a DNA sequence of 229,354 bases using a pseudo random number generator. 
```{r}
set.seed(2017)
# Generate simple random sample
N <- 229354
n <- 296
uniform <- sample.int(N, size = n)
locations <- location$location
```

# Scenario 2: Locations
## Visulization of palindrome location distribution vs Uniform distribution
Using graphical methods to examine if the locations follow uniform distribution.
The pink histogram and the red density curve represents distribution for actual data.
The light blue histogram and the blue density curve represents the distribution and density for the simulated simple random sample.
The purple histogram represents the overlapped area.
```{r}
hist(locations, breaks = 20, probability = TRUE, col = rgb(1,0,0,0.5), main = "Location Data Distribution Comparison", xlab= "Palindrome Locations")
lines(density(locations, adjust = 2), col = 2)
hist(uniform, breaks = 20, probability = TRUE, col = rgb(0,0,1,0.5), add = TRUE)
lines(density(uniform, adjust = 2), col = 4)
pop <- location$location
hist(pop, breaks = 20, probability = TRUE, col = rgb(1,0,0,0.5), main = "Location Data Distribution Comparison", xlab= "Palindrome Locations")
lines(density(pop, adjust = 2), col = 2)
hist(uniform,breaks = 20, probability = TRUE, col = rgb(0,0,1,0.5), add = TRUE)
lines(density(uniform, adjust = 2), col = 4)
legend(x = 180000, y = 0.000008, legend = c("Sample", "Uniform"), lty = c(1,1), col = c(rgb(1,0,0,0.5), rgb(0,0,1,0.5)))
```

In Case 1, we group locations into 50 sub-intervals, and construct chi-square test statistic to compare it against the theoretical uniform distribution.
## Chi-square Goodness of Fit Test
```{r}
k <- 50
locations.expected <- n/k
tab <- table(cut(locations, breaks = seq(0, N, length.out = k+1), include.lowest = TRUE))
locations.observed <- as.vector(tab)
chi_2 <- sum((locations.observed - locations.expected)^2/locations.expected)
chi_2

chi2_compare <- qchisq(p = 0.95, df = 49)
chi2_compare

p_value <- pchisq(chi_2, df = 49, lower.tail = FALSE)
p_value
```


## Visualization of the Residual
```{r}
Residuals <- (locations.observed - locations.expected) / sqrt(locations.expected)
plot(Residuals, type = 'h', ylab = "Standardized Residuals", xlab = "Palindrome locations", main = "Plot of Standardized Residual for Locations")
```

In case 2, we group locations into 100 sub-intervals, and construct chi-square test statistic to compare it against the theoretical uniform distribution.
## Chi-square Goodness of Fit Test
```{r}
k <- 100
locations.expected <- n/k
tab <- table(cut(locations, breaks = seq(0, N, length.out = k+1), include.lowest = TRUE))
locations.observed <- as.vector(tab)
chi_2 <- sum((locations.observed - locations.expected)^2/locations.expected)
chi_2

chi2_compare <- qchisq(p = 0.95, df = 99)
chi2_compare

p_value <- pchisq(chi_2, df = 99, lower.tail = FALSE)
p_value
```
## Visualization of the Residual
```{r}
Residuals <- (locations.observed - locations.expected) / sqrt(locations.expected)
plot(Residuals, type = 'h', ylab = "Standardized Residuals", xlab = "Palindrome locations", main = "Plot of Standardized Residual for Locations")
```

In case 3, we group locations into 500 sub-intervals, and construct chi-square test statistic to compare it against the theoretical uniform distribution.
## Chi-square Goodness of Fit Test
```{r}
k <- 500
locations.expected <- n/k
tab <- table(cut(locations, breaks = seq(0, N, length.out = k+1), include.lowest = TRUE))
locations.observed <- as.vector(tab)
chi_2 <- sum((locations.observed - locations.expected)^2/locations.expected)
chi_2

chi2_compare <- qchisq(p = 0.95, df = 499)
chi2_compare

p_value <- pchisq(chi_2, df = 499, lower.tail = FALSE)
p_value
```
## Visualization of the Residual
```{r}
Residuals <- (locations.observed - locations.expected) / sqrt(locations.expected)
plot(Residuals, type = 'h', ylab = "Standardized Residuals", xlab = "Palindrome locations", main = "Plot of Standardized Residual for Locations")
```

Inference:
Null Hypothesis: Locations are not distributed uniformly.
Conclusion: Since p-value of this chi-square test is smaller than 0.05, we fail to reject the null hypothesis. It indicates that deviations as large as ours are not so likely. Hence, we conclude that it appears that uniform is not a reasonable initial model. 
Residual Conclusion:
Since the value of the standardized residual is larger than 3, the probability model of a uniform distribution is lack of fit.







# Scenario 2: Spacings
# Visulization of palindrome spacings distribution
Using graphical methods to compare a generated Exponential sequence and the distribution of the palindromes counts in different sets of location intervals.
Remark:
The pink histogram and the red density curve represents the distribution and density for actual data.
The light blue histogram and the blue density curve represents distribution for generated Exponential sequence.
The purple histogram represents the overlapped area.
```{r}
locations.sorted = sort(locations, decreasing = FALSE)
distance.pair <- abs(locations.sorted[-1]-locations.sorted[-length(locations)])
distance.pair
```

```{r}
hist(distance.pair, breaks= 15, col = rgb(1,0,0,0.5), probability = TRUE, main = "Consecutive Pairs Spacings Distribution Comparison", xlab = "Distance between Consecutive Palindromes Locations", ylim = c(0,0.001))
lines(density(distance.pair, adjust = 2), col = rgb(1,0,0,0.5))
Expo <- rexp(n-1, rate = 1/mean(distance.pair))
hist(Expo, breaks = 15, col = rgb(0,0,1,0.5), probability = TRUE, add = TRUE)
lines(density(Expo, adjust = 2), col = rgb(0,0,1,0.5))
legend(x = 4200, y = 0.0009, legend = c("Sample", "Exponential"), lty = c(1,1), col = c(rgb(1,0,0,0.5), rgb(0,0,1,0.5)))

```

# Chi-square Goodness of Fit Test (Need to be revised)
We computed that by Maximum likelihood and Method of Moments, the estimator for rate parameter lambda in an Exponential distibution is the 1/ x bar, which is the reciprocal of the average sample rate. This estimator is unbiased.

```{r}
# Construct palindrome count number
# Construct observed number of intervals
spacings.observed <- distance.pair

# Construct expected number of intervals
spacings.expected <- c()
lambda <- 1/mean(distance.pair)
spacings.expected <- rep((n-1)*exp(-lambda),n-1)

chi_2 <- sum((spacings.observed - spacings.expected)^2/spacings.expected)
chi_2

chi2_compare <- qchisq(p = 0.95, df = n-3)
chi2_compare

p_value <- pchisq(chi_2, df = n-3, lower.tail = FALSE)
p_value

# Visualization of the Residual (Need to be revised)

```



# Counts
# Visulization of palindrome counts distribution
Group the data by dividing the interval [0,N] into 61 sub-intervals and count how many data points are inside each of them.
Using graphical methods to compare a generated Poisson sequence and the distribution of the palindromes counts in different sets of location intervals.
Remark:
The pink histogram and the red density curve represents the distribution and density for actual data.
The light blue histogram and the blue density curve represents distribution for generated Poisson sequence.
The purple histogram represents the overlapped area.

```{r}
k <- 61
tab <- table(cut(locations, breaks = seq(0, N, length.out = k+1), include.lowest = TRUE))
counts <- as.vector(tab)
counts

hist(counts, breaks = 15, col = rgb(1,0,0,0.5), probability = TRUE, main = "Counts Distribution Comparison (61 Sub-intervals)", xlab = "Number of Palindromes Sites Inside an Interval", ylim = c(0,0.3))
lines(density(counts, adjust = 2), col = rgb(1,0,0,0.5))
Pois <- rpois(n, lambda = mean(counts))
hist(Pois, breaks = 15, col = rgb(0,0,1,0.5), probability = TRUE, add = TRUE)
lines(density(Pois, adjust = 2), col = rgb(0,0,1,0.5))
legend(x = 13, y = 0.15, legend = c("sample", "Poisson"), lty = c(1,1), col = c(rgb(1,0,0,0.5), rgb(0,0,1,0.5)))
```

# Additional testings on using different lengths of sub-intervals:
```{r}
j <- 1000
tab <- table(cut(locations, breaks = seq(0, N, length.out = j+1), include.lowest = TRUE))
counts <- as.vector(tab)
hist(counts, breaks = 10, col = rgb(1,0,0,0.5), probability = TRUE, main = "Counts Distribution Comparison (1000 Sub-intervals)", xlab = "Number of Palindromes Sites Inside an Interval", ylim = c(0,1))
lines(density(counts, adjust = 2), col = rgb(1,0,0,0.5))
Pois <- rpois(n, lambda = mean(counts))
hist(Pois, breaks = 10, col = rgb(0,0,1,0.5), probability = TRUE, add = TRUE)
lines(density(Pois, adjust = 2), col = rgb(0,0,1,0.5))
legend(x = 32, y = 0.17, legend = c("sample", "Poisson"), lty = c(1,1), col = c(rgb(1,0,0,0.5), rgb(0,0,1,0.5)))
```

```{r}
j <- 45
tab <- table(cut(locations, breaks = seq(0, N, length.out = j+1), include.lowest = TRUE))
counts <- as.vector(tab)
counts
hist(counts, breaks = 15, col = rgb(1,0,0,0.5), probability = TRUE, main = "Counts Distribution Comparison (45 Sub-intervals)", xlab = "Number of Palindromes Sites Inside an Interval", ylim = c(0,0.2))
lines(density(counts, adjust = 2), col = rgb(1,0,0,0.5))
Pois <- rpois(n, lambda = mean(counts))
hist(Pois, breaks = 15, col = rgb(0,0,1,0.5), probability = TRUE, add = TRUE)
lines(density(Pois, adjust = 2), col = rgb(0,0,1,0.5))
legend(x = 20, y = 0.15, legend = c("sample", "Poisson"), lty = c(1,1), col = c(rgb(1,0,0,0.5), rgb(0,0,1,0.5)))
```

```{r}
j <- 27
tab <- table(cut(locations, breaks = seq(0, N, length.out = j+1), include.lowest = TRUE))
tab
counts <- as.vector(tab)
counts
hist(counts, breaks = 10, col = rgb(1,0,0,0.5), probability = TRUE, main = "Counts Distribution Comparison (27 Sub-intervals)", xlab = "Number of Palindromes Sites Inside an Interval", ylim = c(0,0.2))
lines(density(counts, adjust = 2), col = rgb(1,0,0,0.5))
Pois <- rpois(n, lambda = mean(counts))
hist(Pois, breaks = 10, col = rgb(0,0,1,0.5), probability = TRUE, add = TRUE)
lines(density(Pois, adjust = 2), col = rgb(0,0,1,0.5))
legend(x = 18, y = 0.17, legend = c("sample", "Poisson"), lty = c(1,1), col = c(rgb(1,0,0,0.5), rgb(0,0,1,0.5)))
```

# Chi-square Goodness of Fit Test
We computed that by Maximum likelihood and Method of Moments, the estimator for rate parameter lambda in a Poisson distibution is the x bar, which is the average sample rate. This estimator is unbiased. Since we estimated one parameter, the degree of freedom equals k-1-1 = 59.

Case 1: When number of subintervals k = 61
```{r}
# Divide the CMV DNA into 61 non overlapping regions of length 3583 bases
k <- 61
tab <- table(cut(locations, breaks = seq(0, N, length.out = k+1), include.lowest = TRUE))
counts.obs <- as.vector(tab)
a <- table(cut(counts.obs, breaks = seq(min(counts.obs), max(counts.obs), length.out = k+1), include.lowest = TRUE))
a

# Construct palindrome count number
palindrome.count <- c("0,1,2", "3","4","5","6","7","8+")

# Construct observed number of intervals
intervals.observed <- c(8, 9, 13, 10, 8, 8, 5) 

# Construct expected number of intervals
expected <- c()
lambda <- n/k
for (i in c(0:17)){
  expect <- k* exp(-lambda)* lambda**i /factorial(i)
  expected <- c(expected, expect)
}
sum <- 0
for (j in c(9:17)){
  sum <- sum+expected[j]
}
intervals.expected <- c(expected[1]+expected[2]+expected[3],expected[4],expected[5],expected[6],expected[7],expected[8],sum)

# Create contingency table
b <- data.frame(palindrome.count,intervals.observed,intervals.expected)
b

chi_2 <- sum((intervals.observed - intervals.expected)^2/intervals.expected)
chi_2

chi2_compare <- qchisq(p = 0.95, df = 7)
chi2_compare

p_value <- pchisq(chi_2, df = 7, lower.tail = FALSE)
p_value

```

# Visualization of the Residual
```{r}
Residuals <- (intervals.observed-intervals.expected) / sqrt(intervals.expected)
plot(Residuals, type = 'h', ylab = "Standardized Residuals", xlab = "Palindrome Counts", main = "Plot of Standardized Residual for Counts")
```

Case 1: When number of subintervals k = 1000
```{r}
# Divide the CMV DNA into 61 non overlapping regions of length 3583 bases
k <- 1000
tab <- table(cut(locations, breaks = seq(0, N, length.out = k+1), include.lowest = TRUE))
counts.obs <- as.vector(tab)
a <- table(cut(counts.obs, breaks = seq(min(counts.obs), max(counts.obs), length.out = k+1), include.lowest = TRUE))
a



```
```{r}
# Construct palindrome count number
palindrome.count <- c("0,1,2", "3","4","5","6","7","8+")

# Construct observed number of intervals
intervals.observed <- c(8, 9, 13, 10, 8, 8, 5) 

# Construct expected number of intervals
expected <- c()
lambda <- n/k
for (i in c(0:17)){
  expect <- k* exp(-lambda)* lambda**i /factorial(i)
  expected <- c(expected, expect)
}
sum <- 0
for (j in c(9:17)){
  sum <- sum+expected[j]
}
intervals.expected <- c(expected[1]+expected[2]+expected[3],expected[4],expected[5],expected[6],expected[7],expected[8],sum)

# Create contingency table
b <- data.frame(palindrome.count,intervals.observed,intervals.expected)
b

chi_2 <- sum((intervals.observed - intervals.expected)^2/intervals.expected)
chi_2

chi2_compare <- qchisq(p = 0.95, df = 7)
chi2_compare

p_value <- pchisq(chi_2, df = 7, lower.tail = FALSE)
p_value

```



Null Hypothesis: Counts are not distributed according to Poisson distribution.
Conclusion: Since p-value of this chi-square test is 1, we will reject the null hypothesis. From this computation, we see that deviations as large as ours are very much likely to be observed. Hence, we conclude that it appears that Poisson is a reasonable initial model. 



