---
title: "Math189Pr_3"
output: html_document
---

Overall Question: How do we find clusters of palindromes? How do we determine whether a cluster is just a chance occurrence or a potential replication site?

# Import Dataset
```{r}
location <- read.table("hcmv-25kgjn1-1rfrtkc.txt", header = TRUE)
raw <- read.csv("RAW_DATA-2iwcznn-2kr2xw0.csv", header = TRUE)
```

# Random Scatter
Generate structure that examine the locations of the palindromes. Simulate 296 palindrome sites chosen at random along a DNA sequence of 229,354 bases using a pseudo random number generator. 
```{r}
set.seed(2017)
# Generate simple random sample
N <- 229354
n <- 296
pop <- sample.int( N, size = n)
```

# Visulization of palindrome location distribution
Using graphical methods to examine if the locations follow uniform distribution.
The yellow histogram and the blue density curve represents the distribution and density for generated simple random sample.
The green histogram and the red density curve represents distribution for actual data.
The purple histogram represents the overlapped area.
```{r}
hist(pop, breaks = 100, probability = TRUE, col = 7, main = "Location Data Distribution Comparison", xlab= "Palindrome Locations")
lines(density(pop, adjust = 2), col = 4)
hist(location$location, breaks = 100, probability = TRUE, col = 3, add = TRUE)
lines(density(location$location, adjust = 2), col = 2)
```
Conclusion:
Graphically, the distribution of the locations of palindromes approximately follows uniform distribution. 

# Visulization of palindrome counts distribution
Group the data by dividing the interval [0,230000] into 45 sub-intervals and count how many data points are inside each of them.
```{r}
k <- 45
tab <- table(cut(pop, breaks = seq(0, 230000, length.out = k+1), include.lowest = TRUE))
counts <- as.vector(tab)
```
Using graphical methods to compare a generated Poisson(Theoretical) sequence and the distribution of the palindromes counts in bins of locations.
Remark:
The pink histogram and the red density curve represents the distribution and density for actual data.
The light blue histogram and the blue density curve represents distribution for generated Poisson sequence.
The purple histogram represents the overlapped area.
```{r}
hist(counts, breaks = 15, col = rgb(1,0,0,0.5), probability = TRUE, main = "Counts Distribution Comparison (Sub-interval length = 45)", xlab = "Number of Palindromes Sites Inside an Interval", ylim = c(0,0.2))
lines(density(counts, adjust = 2), col = rgb(1,0,0,0.5))
Pois <- rpois(1000, lambda = mean(counts))
hist(Pois, breaks = 15, col = rgb(0,0,1,0.5), probability = TRUE, add = TRUE)
lines(density(Pois, adjust = 2), col = rgb(0,0,1,0.5))
legend(x = 14, y = 0.15, legend = c("sample", "Poisson"), lty = c(1,1), col = c(rgb(1,0,0,0.5), rgb(0,0,1,0.5)))
```

# Additional testings on using different lengths of sub-intervals:
```{r}
j <- 10
tab <- table(cut(pop, breaks = seq(0, 230000, length.out = j+1), include.lowest = TRUE))
tab
counts <- as.vector(tab)
counts

hist(counts, breaks = 15, col = rgb(1,0,0,0.5), probability = TRUE, main = "Counts Distribution Comparison (Sub-interval length = 10)", xlab = "Number of Palindromes Sites Inside an Interval", ylim = c(0,0.2))
lines(density(counts, adjust = 2), col = rgb(1,0,0,0.5))
Pois <- rpois(1000, lambda = mean(counts))
hist(Pois, breaks = 15, col = rgb(0,0,1,0.5), probability = TRUE, add = TRUE)
lines(density(Pois, adjust = 2), col = rgb(0,0,1,0.5))
legend(x = 14, y = 0.15, legend = c("sample", "Poisson"), lty = c(1,1), col = c(rgb(1,0,0,0.5), rgb(0,0,1,0.5)))
```

```{r}
i <- 100
tab <- table(cut(pop, breaks = seq(0, 230000, length.out = i+1), include.lowest = TRUE))
tab
counts <- as.vector(tab)
counts
hist(counts, breaks = 15, col = rgb(1,0,0,0.5), probability = TRUE, main = "Counts Distribution Comparison (Sub-interval length = 100)", xlab = "Number of Palindromes Sites Inside an Interval", ylim = c(0,0.55))
lines(density(counts, adjust = 2), col = rgb(1,0,0,0.5))
Pois <- rpois(1000, lambda = mean(counts))
hist(Pois, breaks = 15, col = rgb(0,0,1,0.5), probability = TRUE, add = TRUE)
lines(density(Pois, adjust = 2), col = rgb(0,0,1,0.5))
legend(x = 14, y = 0.15, legend = c("sample", "Poisson"), lty = c(1,1), col = c(rgb(1,0,0,0.5), rgb(0,0,1,0.5)))
```

```{r}
d<- 1000
tab <- table(cut(pop, breaks = seq(0, 230000, length.out = d+1), include.lowest = TRUE))
tab
counts <- as.vector(tab)
counts
hist(counts, breaks = 15, col = rgb(1,0,0,0.5), probability = TRUE, main = "Counts Distribution Comparison (Sub-interval length = 1000)", xlab = "Number of Palindromes Sites Inside an Interval", ylim = c(0,1))
lines(density(counts, adjust = 2), col = rgb(1,0,0,0.5))
Pois <- rpois(1000, lambda = mean(counts))
hist(Pois, breaks = 15, col = rgb(0,0,1,0.5), probability = TRUE, add = TRUE)
lines(density(Pois, adjust = 2), col = rgb(0,0,1,0.5))
legend(x = 14, y = 0.15, legend = c("sample", "Poisson"), lty = c(1,1), col = c(rgb(1,0,0,0.5), rgb(0,0,1,0.5)))
```
Conclusion: 
Since the shape of the distibutions are nearly the same while the length of sub-intervals is changing, the distribution of palindromes counts approximately follows Poisson Distribution graphically.

# Visulization of palindrome spacings distribution
```{r}
pop1 = sort(pop, decreasing = FALSE)
pop.shifted <- pop1[-1]
pop.original <- pop1[-length(pop)]
distance <- pop.shifted-pop.original
distance <- abs(distance)
```


```{r}
hist(distance, breaks= 15, col = rgb(1,0,0,0.5), probability = TRUE, main = "Spacings Distribution Comparison", xlab = "Distance between Consecutive Palindromes Locations", ylim = c(0,0.001))
lines(density(distance, adjust = 2), col = rgb(1,0,0,0.5))
Expo <- rexp(1000, rate = 1/mean(distance))
hist(Expo, breaks = 15, col = rgb(0,0,1,0.5), probability = TRUE, add = TRUE)
lines(density(Expo, adjust = 2), col = rgb(0,0,1,0.5))
legend(x = 14, y = 0.15, legend = c("sample", "Poisson"), lty = c(1,1), col = c(rgb(1,0,0,0.5), rgb(0,0,1,0.5)))

```

# Chi-square Goodness of Fit Test
Testing for Counts
Group the data by dividing the interval [0,230000] into 45 sub-intervals and count how many data points are inside each of them.
```{r}
k <- 45
counts.theo <- n/k
tab <- table(cut(pop, breaks = seq(0, 230000, length.out = k+1), include.lowest = TRUE))
counts.obs <- as.vector(tab)
chi_2 <- sum((counts.obs - counts.theo)^2/counts.theo)
chi_2

chi2_compare <- qchisq(p = 0.95, df = 44)
chi2_compare

p_value <- pchisq(chi_2, df = 44, lower.tail = FALSE)
p_value
```

# Visualization of the Residual
```{r}
Residuals <- (counts.obs - counts.theo) / sqrt(counts.theo)
plot(Residuals, type = 'h', ylab = "Standardized Residuals", xlab = "Palindrome Counts", main = "Plot of Standardized Residual")
```

Testing for spacings
```{r}
# Need to be revised
k <- 45
spacings.theo <- N/k
spacings.obs <- distance
chi_2 <- sum((spacings.obs - spacings.theo)^2/spacings.theo)
chi_2

chi2_compare <- qchisq(p = 0.95, df = 44)
chi2_compare

p_value <- pchisq(chi_2, df = 44, lower.tail = FALSE)
p_value
```

# Design Test Statistic Testing for the Maximum Cluster
```{r}

```

# Additional Question and Hypothesis
```{r}

```


